Case Study: Gene Set Enrichment Analysis (Mouse Data Set)

### 2018-07-23 datasets/GSEA/gsea-human.Rmd
Add filter to GSEA multiple testing

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)
```

@@

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
    strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=18)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
    strat_hist( pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
```

@@

We exclude `ashr`, `scott-theoretical` and `scott-empirical` from the analysis because
the necessary assumptions are not met in the current case study. Namely, effect sizes
and standard errors are not available for ASH, and the test statistics are not normally
distributed under the null and alternative, as required by Scott's FDR regression methods.

>>>>

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)
sum(genes, na.rm=TRUE)
```

@@

We filter out really small gene sets and those gene sets which have no DE genes
in them (many of these are set to 1).

```{r}
goRes <- goRes %>%
  dplyr::filter(numDEInCat > 0, numInCat > 10 )
```

@@

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="numInCat", maxy=7)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
```

@@

We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
the necessary assumptions are not met in the current case study. Namely, effect sizes
and standard errors are not available for ASH, and the test statistics are not normally
distributed under the null and alternative, as required by Scott's FDR regression methods.

### 2018-07-23 datasets/GSEA/gsea-mouse.Rmd
Add filter to GSEA multiple testing

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%
    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=35)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
    strat_hist( pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
```

>>>>

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval = "over_represented_pvalue", covariate = "numInCat", maxy=7)
```

@@

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
```
